A Large-Scale Study of the Use of Eval in R

### 2020-08-28 inst/paper/main.tex
Add explanation for CRAN callsites

The most frequent ten expressions to eval calls from CRAN packages contribute to
90\% of all \eval calls.
\begin{table*}[tb]
  \centering
  \begin{tabular}{l|rr} \hline
    Expression & Eval Call & Eval Call \% \\\hline
    \c{\<Environment\>} &                                                          1091474  & 68.2\%\\
    \c{column[rows] <<- what} &                                                    55677    & 3.5\%\\
    \c{function(value) freduce(value, `_function_list`)} &                         37251    & 2.3\%\\
    \c{c("default", "default2012", "default2011" ...}&                             20610    & 1.3\%\\
    \c{force(..1)}                                        &                        20461    & 1.3\%\\
    \c{alist(`_spec`)}                                   &                         18532    & 1.2\%\\
    \c{inner}                                           &                          18530    & 1.2\%\\
    \c{slot(family, "fini")}                           &                           7288     & 0.5\%\\
    \c{\~fitted(.)}                                    &                           4733     & 0.3\%\\
    \c{function(.) purrr::keep(., is.function)}      &                             4344     & 0.3\%\\\hline
  \end{tabular}
  \caption{Top ten eval calls in CRAN}
  \label{A}
\end{table*}
The most frequently occurring call is to the
\eval in \c{ggplot2::ggproto} which happens over 68\% of the times. The
remaining expressions occur with a frequency of less than 4\%. They originate
from \c{plyr}, \c{VGAM}, \c{magrittr}, \c{??}

>>>>

The most frequent ten expressions to eval calls from CRAN packages contribute to
77.1\% of all \eval calls.
\begin{table*}[tb]
  \centering
  \begin{tabular}{l|rr} \hline
    Expression & Eval Call & Eval Call \% \\\hline
    \c{Environment} &                                                              989302   & 61.8\%\\
    \c{column[rows] <<- what} &                                                    55677    & 3.5\%\\
    \c{function(value) freduce(value, `_function_list`)} &                         37251    & 2.3\%\\
    \c{NULL} &                         32005    & 2\%\\
    \c{List} &                         22293    & 1.4\%\\
    \c{c("default", "default2012", "default2011" ...}&                             20610    & 1.3\%\\
    \c{force(..1)}                                        &                        20461    & 1.3\%\\
    \c{alist(`_spec`)}                                   &                         18532    & 1.2\%\\
    \c{inner}                                           &                          18530    & 1.2\%\\
    \c{String Vector}                           &                           17487     & 1.1\%\\
  \end{tabular}
  \caption{Top ten eval calls in CRAN}
  \label{A}
\end{table*}

The expression \c{Environment} occurs because of the four callsites explained
above, \c{ggplot2::ggproto}, \c{R6::generator_funs}, \c{future::backtrace} and
\c{RModel::str.RMmodel}. The next expression, \c{column[rows] <<- what}, is used
inside the \c{plyr::rbind.fill} function to merge data frames by assigning
concatenated vectors to rows. The \c{<<-} operator is interesting in that it
skips the current scope and assigns in a parent scope in which the variable is
already present. In our corpus, all these \eval calls contribute to a single
side-effect. The expression \c{function(value) freduce(value, `_function_list`)}
arises from the \c{magrittr::\%>\%} function which is a pipe operator that pipes
the output of previous command to the next one. The expression is evaluated in a
custom environment to create a function binding for evaluating the components of
the pipe.\c{String Vector} and \c{List} also arise from the same function when a
string or a list is piped using the \c{\%>\%} function into the next expression.
The \c{NULL} arises from \c{R6::generator_funs} function when the \eval is
passed a \c{NULL} argument by the \c{DataMask_generator} package. The
\c{c("default", "default2012", "default2011" ...} pattern arises from
\c{copula::polyG} where it reflectively access the default expression for its
formal parameter and evaluates it. The \c{force(..1)} and \c{alist(`_spec`)}
patterns occur in \c{glue::glue_data} function which concatenates and
interpolates strings. The two patterns occur because the function captures
unevaluated unnamed arguments and maps the evaluation of \c{force(..1)} on them.
The \c{force} function forces promises~\cite{oopsla19a} and returns the result
of evaluation. The \c{inner} pattern arises from \c{glue::identity_transformer}
which enables the creation of custom transformation functions for affecting the
interpolation and concatenation of input by the \c{glue} package.
